### 2018-06-18 CS761-RandomizedAlgorithms/assignments/a2.tex
wrote up 5

	In the inductive step, consider an edge $e=v_iv_j$. Let $a_1 \dots a_\ell$ be the global encoding vectors of incoming edges to $v_i$. Then each $a_i$ is a polynomial in the local encoding coefficients of degree at most $i-2$ by induction. Now, the global encoding vector for $e$ is
	$$b_1^{(e)}a_1 + \dots + b_\ell^{(e)}a_\ell$$
	which is a sum of multivariate polynomials of degree at most $i-1$, and hence is a polynomial of degree at most $i-1$ as desired.
\end{proof}
\paragraph{}
For vertex $t \in V\backslash \{s\}$ with $\ell$ incoming arcs, let $(c_{i1}, \dots, c_{id})$ be the global encoding vector of the $i$-th incoming edge to $t$. Then we can define the $\ell \times d$ receiver matrix $C$, whose entries are multivariate polynomails over $\F$, where entry $C_{ij} = c_{ij}$ for all $i,j$.
\paragraph{}
Let $C'$ be the largest non-singular submatrix of $C$. The size of $C'$ corresponds to the rank of $C$. So if $\det(C')$ is not $0$ after assigning values to its entries, then the rank of $C$ will be preserved. Since $\det(C')$ is not identically $0$, by Schwartz-Zippel the probability that $\det(C') =0$ after assigning random values for the local encoding coefficients is at most 
$$\frac{dn}{|\F|}$$
Since our Claim implies that the entries of $C'$ are each of degree at most $n$ and so $\det(C')$ is a polynomial of degree at most $dn$. Since $d\leq n$, if we take $|\F| \geq n^4$ this probability of failure is at most 
$$\frac{1}{n^2}.$$
So by the union bound, the probability that the rank of the receiver matrix for each vertex is preserved is at least
$$1-\frac{1}{n}.$$
\paragraph{}
It remains to show that the rank of $C$ is equal to the edge connectivity of $t$. Suppose the edge-connectivity of $t$ is $k$. On the one hand there are $k$ edge-disjoint paths to $t$. If we set the message on each edge of the $i$-th such path to be $e_i$ and the message on every other edge to be $0$ then $C$ will contain (possibly up to permuting the rows) the size $k$ identity matrix, and hence as a matrix of polynomials, the rank of $C$ is at least $k$. On the other hand, by the max-flow min-cut theorem there is an $s$-$t$ cut with exactly $k$ edges. So every row of $C$ is a linear combination of the $k$ vectors along that cut. Hence the rank of $C$ is at most $k$. Therefore the rank of $C$ is equal to $k$. So we have shown the edge-connectivity algorithm works with high probability when you choose a finite field $\F$ so that $|F| \geq n^4 = \Theta(\text{poly}(n))$. $\blacksquare$
\paragraph{Implementation}
We now present an efficient implementation.
\end{document}

>>>>

	In the inductive step, consider an edge $e=v_iv_j$. Let $a_1 \dots a_\ell$ be the global encoding vectors of incoming edges to $v_i$. Then each $a_i$ is a polynomial in the local encoding coefficients of degree at most $i-2$ by induction. Now, the global encoding vector for $e$ is
	$$b_1^{(e)}a_1 + \dots + b_\ell^{(e)}a_\ell$$
	which is a sum of multivariate polynomials of degree at most $i-1$, and hence is a polynomial of degree at most $i-1$ as desired.
\end{proof}
\paragraph{}
For vertex $t \in V\backslash \{s\}$ with $\ell$ incoming arcs, let $(c_{i1}, \dots, c_{id})$ be the global encoding vector of the $i$-th incoming edge to $t$. Then we can define the $\ell \times d$ receiver matrix $C$, whose entries are multivariate polynomails over $\F$, where entry $C_{ij} = c_{ij}$ for all $i,j$.
\paragraph{}
Let $C'$ be the largest non-singular submatrix of $C$. The size of $C'$ corresponds to the rank of $C$. So if $\det(C')$ is not $0$ after assigning values to its entries, then the rank of $C$ will be preserved. Since $\det(C')$ is not identically $0$, by Schwartz-Zippel the probability that $\det(C') =0$ after assigning random values for the local encoding coefficients is at most 
$$\frac{dn}{|\F|}$$
Since our Claim implies that the entries of $C'$ are each of degree at most $n$ and so $\det(C')$ is a polynomial of degree at most $dn$. Since $d\leq n$, if we take $|\F| \geq n^4$ this probability of failure is at most 
$$\frac{1}{n^2}.$$
So by the union bound, the probability that the rank of the receiver matrix for each vertex is preserved is at least
$$1-\frac{1}{n}.$$
\paragraph{}
It remains to show that the rank of $C$ is equal to the edge connectivity of $t$. Suppose the edge-connectivity of $t$ is $k$. On the one hand there are $k$ edge-disjoint paths to $t$. If we set the message on each edge of the $i$-th such path to be $e_i$ and the message on every other edge to be $0$ then $C$ will contain (possibly up to permuting the rows) the size $k$ identity matrix, and hence as a matrix of polynomials, the rank of $C$ is at least $k$. On the other hand, by the max-flow min-cut theorem there is an $s$-$t$ cut with exactly $k$ edges. So every row of $C$ is a linear combination of the $k$ vectors along that cut. Hence the rank of $C$ is at most $k$. Therefore the rank of $C$ is equal to $k$. So we have shown the edge-connectivity algorithm works with high probability when you choose a finite field $\F$ so that $|F| \geq n^4 = \Theta(\text{poly}(n))$. $\blacksquare$
\paragraph{Implementation}
We now present an efficient implementation. For each vertex $v \in V$ we replace $v$ with a super-concentrator of size
$$O(\text{in-degree}(v) + \text{out-degree}(v)\}) = O(\deg(v))$$
where $\deg(v)$ is the number of edges incident to $v$ in the underlying undirected graph to $G$. Since the size of the super-concentrator we replace each vertex with is an upper-bound on the number of edge-disjoint paths in $G$ which can use that vertex, we preserve edge-connectivity after this transformation. From the notes, the encoding at each vertex can be done in $O(d)$ time after the transformation. The number of vertices in the new graph is 
$$\sum_{v \in V} O(\deg(v)) = O(2|E|) = O(|E|)$$
by the handshaking-lemma. So the entire encoding can be done in $O(d|E|)$ time.
\paragraph{}
Using the random matrix rank algorithm for Lecture $9$ we can compute the rank of the matrix associated with each vertex in time
$$O(d\cdot \deg(v) + \min\{d,\deg(v)\}^\omega)$$
where $\omega$ is the matrix multiplication constant, with high probability. Summing over vertices $v \in V$, we can compute all the edge connectivities in time
$$O(d|E| + |V|d^\omega).$$
Thus our implementation will succeed in computing the edge-connectivity from $s$ to $t$ for each $t \in V$ with high probability, in time
$$O(d|E| +|V|d^{\omega}).$$
$\blacksquare$
\end{document}
